Sorghum
========================================================

Set up

```{r fig.width=10, fig.height=6}
library(ggplot2)
# normalisation function
normalise <- function(df) {
  df$count <- 1000000*(df$count/sum(df$count))
  return(df)
}
```

### Raw Data

```{r fig.width=10, fig.height=6}
setwd("~/hydrogen/sorghum")
fastq <- read.table("raw_read_counts.txt", sep="\t", header=FALSE)
names(fastq) <- c("count", "cell", "rep") # changing header names

ggplot(fastq, aes(x=paste(cell,rep), y=count, fill=cell)) +
  geom_bar(stat="identity") +
  ggtitle("Raw read counts") +
  geom_text(aes(label=count, y=count+400000)) +
  labs(x = "Cell and rep") +
  ylim(0,30000000)

```

Reads trimmed with trimmomatic

```{r fig.width=10, fig.height=6}
setwd("~/hydrogen/sorghum")
fastq <- read.table("t_read_counts.txt", sep="\t", header=FALSE)
names(fastq) <- c("count", "cell", "rep") # changing header names

ggplot(fastq, aes(x=paste(cell,rep), y=count, fill=cell)) +
  geom_bar(stat="identity") +
  ggtitle("Raw read counts") +
  geom_text(aes(label=count, y=count+400000)) +
  labs(x = "Cell and rep") +
  ylim(0,30000000)

```

Reads trimmed with fastq-mcf

```{r fig.width=10, fig.height=6}
setwd("~/hydrogen/sorghum")
fastq <- read.table("mcf_read_counts.txt", sep="\t", header=FALSE)
names(fastq) <- c("count", "cell", "rep") # changing header names

ggplot(fastq, aes(x=paste(cell,rep), y=count, fill=cell)) +
  geom_bar(stat="identity") +
  ggtitle("Raw read counts") +
  geom_text(aes(label=count, y=count+400000)) +
  labs(x = "Cell and rep") +
  ylim(0,30000000)

```

### Read lengths

Bundle Sheath (with Trimmomatic)

```{r fig.width=10, fig.height=6}
setwd("~/hydrogen/sorghum")
bs_data <- data.frame()
for (i in 1:3) {
  tmp <- read.table(paste("t_read_length_BS-",i,".txt",sep=""), header=FALSE)
  names(tmp) <- c("length", "count") # changing header names
  tmp$rep <- paste("rep",i, sep="-")
  tmp <- normalise(tmp)
  bs_data <- rbind(bs_data,tmp)
}

ggplot(bs_data, aes(x=length, y=count, colour=rep)) +
  geom_line(size=1) +
  scale_colour_brewer(type="qual", palette=6) +
  ggtitle("Bundle Sheath sRNA read lengths (trimmomatic)") +
  xlim(17,49)

```

Bundle Sheath (with fastq-mcf)

```{r fig.width=10, fig.height=6}
setwd("~/hydrogen/sorghum")
bs_data <- data.frame()
for (i in 1:3) {
  tmp <- read.table(paste("mcf_read_length_BS-",i,".txt",sep=""), header=FALSE)
  names(tmp) <- c("length", "count") # changing header names
  tmp$rep <- paste("rep",i, sep="-")
  tmp <- normalise(tmp)
  bs_data <- rbind(bs_data,tmp)
}

ggplot(bs_data, aes(x=length, y=count, colour=rep)) +
  geom_line(size=1) +
  scale_colour_brewer(type="qual", palette=6) +
  ggtitle("Bundle Sheath sRNA read lengths (fastq-mcf)") +
  xlim(17,49)

```


Mesophyll (with Trimmomatic)

```{r fig.width=10, fig.height=6}
setwd("~/hydrogen/sorghum")
bs_data <- data.frame()
for (i in 1:3) {
  tmp <- read.table(paste("t_read_length_M-",i,".txt",sep=""), header=FALSE)
  names(tmp) <- c("length", "count") # changing header names
  tmp$rep <- paste("rep",i, sep="-")
  tmp <- normalise(tmp)
  bs_data <- rbind(bs_data,tmp)
}

ggplot(bs_data, aes(x=length, y=count, colour=rep)) +
  geom_line(size=1) +
  scale_colour_brewer(type="qual", palette=6) +
  ggtitle("Mesophyll sRNA read lengths (trimmomatic)") +
  xlim(17,49)

```

Bundle Sheath (with fastq-mcf)

```{r fig.width=10, fig.height=6}
setwd("~/hydrogen/sorghum")
bs_data <- data.frame()
for (i in 1:3) {
  tmp <- read.table(paste("mcf_read_length_M-",i,".txt",sep=""), header=FALSE)
  names(tmp) <- c("length", "count") # changing header names
  tmp$rep <- paste("rep",i, sep="-")
  tmp <- normalise(tmp)
  bs_data <- rbind(bs_data,tmp) 
}

ggplot(bs_data, aes(x=length, y=count, colour=rep)) +
  geom_line(size=1) +
  scale_colour_brewer(type="qual", palette=6) +
  ggtitle("Mesophyll sRNA read lengths (fastq-mcf)") +
  xlim(17,49)

```

### sRNA Expression

Bundle Sheath

```{r fig.width=10, fig.height=6}
# bs_counts <- read.table("output_BS-1_counts.txt", header=FALSE)
# names(bs_counts) <- c("name", "count")
# bs_counts$rep <- paste("rep",1,sep="-")
# 
# ggplot(bs_counts, aes(x = log(count))) + geom_histogram(binwidth=0.5) + scale_y_log10()
```

Mesophyll

```{r fig.width=10, fig.height=6}
# m_counts <- read.table("output_M-2_counts.txt", header=FALSE)
# names(m_counts) <- c("name", "count")
# m_counts$rep <- paste("rep",2,sep="-")
# 
# ggplot(m_counts, aes(x = log(count))) + geom_histogram(binwidth=0.5) + scale_y_log10()
```

Scatter!

```{r fig.width=10, fig.height=6}
# bs_1 <- read.table("output_BS-1_counts.txt", header=FALSE)
# bs_2 <- read.table("output_BS-2_counts.txt", header=FALSE)
# bs_3 <- read.table("output_BS-3_counts.txt", header=FALSE)
# mes_1 <- read.table("output_M-1_counts.txt", header=FALSE)
# mes_2 <- read.table("output_M-2_counts.txt", header=FALSE)
# mes_3 <- read.table("output_M-3_counts.txt", header=FALSE)
# names(bs_1) <- c("name", "count")
# names(bs_2) <- c("name", "count")
# names(bs_3) <- c("name", "count")
# names(mes_1) <- c("name", "count")
# names(mes_2) <- c("name", "count")
# names(mes_3) <- c("name", "count")
# 
# bs_1 <- normalise(bs_1)
# bs_2 <- normalise(bs_2)
# bs_3 <- normalise(bs_3)
# mes_1 <- normalise(mes_1)
# mes_2 <- normalise(mes_2)
# mes_3 <- normalise(mes_3)
```

Compare Rep 1 and 2 in Bundle Sheath

```{r fig.width=10, fig.height=6}
# both <- merge(bs_1, bs_2, by.x = "name", by.y = "name")
# 
# names(both) <- c("name", "BS", "M")
# 
# ggplot(both, aes(x = log(BS), y = log(M))) + geom_point(size = 0.1) + ggtitle("Bundle Sheath - Rep 1 and 2")

```

Compare Rep 2 and 3 in Mesophyll

```{r fig.width=10, fig.height=6}
# both <- merge(mes_2, mes_3, by.x = "name", by.y = "name")
# 
# names(both) <- c("name", "BS", "M")
# 
# ggplot(both, aes(x = log(BS), y = log(M))) + geom_point(size = 0.1) + ggtitle("Mesophyll - Rep 2 and 3")

```

Compare Rep 2 and 3 in Bundle Sheath

```{r fig.width=10, fig.height=6}
# both <- merge(bs_2, bs_3, by.x = "name", by.y = "name")
# 
# names(both) <- c("name", "BS", "M")
# 
# ggplot(both, aes(x = log(BS), y = log(M))) + geom_point(size = 0.1) + ggtitle("Bundle Sheath - Rep 2 and 3")

```

Compare Rep 1 and 2 in Mesophyll

```{r fig.width=10, fig.height=6}
# both <- merge(bs_3, mes_1, by.x = "name", by.y = "name")
# 
# names(both) <- c("name", "BS", "M")
# 
# ggplot(both, aes(x = log(BS), y = log(M))) + geom_point(size = 0.1) + ggtitle("Bundle Sheath 3 - Mesophyll Rep 1")
# 
```

Compare Rep 1 of Bundle Sheath and 2 in Mesophyll

```{r fig.width=10, fig.height=6}
# both <- merge(bs_1, mes_2, by.x = "name", by.y = "name")
# 
# names(both) <- c("name", "BS", "M")
# 
# ggplot(both, aes(x = log(BS), y = log(M))) + geom_point(size = 0.1) + ggtitle("Bundle Sheath rep 1 and Mesophyll rep 2")
# 
```

Correlation Plot

```{r fig.width=10, fig.height=6}
# library("reshape2")
# 
# a <- merge(bs_1, bs_2, by="name")
# bs_all <- merge(a, bs_3, by="name")
# 
# b <- merge(mes_1, mes_2, by="name")
# mes_all <- merge(b, mes_3, by="name")
# 
# bs_m <- merge(bs_all, mes_all, by="name")
# names(bs_m) <- c("names", "bs_1", "bs_2", "bs_3", "m_1", "m_2", "m_3")
# 
# bs_m_c <- cor(bs_m[,2:7])
# 
# melted <- melt(bs_m_c)
# 
# p <- ggplot(data = melted, aes_string(x = names(melted)[1], y = names(melted)[2], fill = "value")) + 
#   geom_tile() + 
#   geom_text(aes(label = round(value, 4))) + xlab("") + ylab("") +
#   ggtitle("Expression Correlation")
# p
# 
```
